---
title: "Predictive Analytics of French CPI (ARIMA)"
output: html_document
---
# Import Package
```{r}
library(tidyverse)
library(lubridate)
library(tsibble)
library(forecast)
library(tseries)
library(rstudioapi)
library(ggplot2)
library(slider)
library(dplyr)
library(feasts)
library(gridExtra)
library(urca)
library(fable)
# Load the data
fcpi <- read.csv("fcpi.csv",sep = ';')
head(fcpi)
# Convert date format
fcpi$time <- as.Date(fcpi$time, format="%Y-%m-%d")
# Convert to yearmonth for tsibble format
fcpi_tsibble <- fcpi %>%
  mutate(time = yearmonth(time)) %>%
  as_tsibble(index = time)
glimpse(fcpi_tsibble)
```
# 1. Plot the data and the autocorrelation function. 
```{r}
# Plot the time series
autoplot(fcpi_tsibble)+
  ggtitle('French CPI(1960-2019)')+
  xlab("Time") + ylab("CPI")
# Plot rolling standard deviation
fcpi_tsibble <- fcpi_tsibble %>%
  mutate(rolling_sd = slide_dbl(cpi, sd, .before = 11, .complete = TRUE))

ggplot(fcpi_tsibble, aes(x = time, y = rolling_sd)) +
  geom_line(color = "blue") +
  labs(title = "Rolling Standard Deviation (12 months)", x = "Time", y = "Standard Deviation") +
  theme_minimal()
```
# 2. Based on the above plot, the variance of the series appears to be heteroscedastic, suggesting that a log transformation is necessary to stabilize the variance.
```{r}
# Logarithm
fcpi_tsibble <- fcpi_tsibble %>%
  mutate(fcpi_log=log(cpi))
# Plot the log time series
autoplot(fcpi_tsibble,fcpi_log)+
  ggtitle('Log-transformed French CPI')+
  xlab('Time')+ylab('Log(CPI)')
```
# 3. This series exibits a clear upward trend and is non-stationary, indicating the need for differencing (first differencing).

```{r fig.width=14, fig.height=5}
# First-Differencing
fcpi_tsibble <- fcpi_tsibble %>%
  mutate(fcpi_log_diff = difference(fcpi_log, differences = 1))

# Plot differenced time series
autoplot(fcpi_tsibble, fcpi_log_diff) +
  ggtitle("Differenced Log CPI (First Difference)") +
  xlab("Time") + ylab("Differenced log(CPI)")

# Plot ACF and PACF
acf_diff_plot <- ggAcf(fcpi_tsibble$fcpi_log_diff) +
  ggtitle("ACF of Differenced Log CPI") 

pacf_diff_plot <- ggPacf(fcpi_tsibble$fcpi_log_diff) +
  ggtitle("PACF of Differenced Log CPI")

grid.arrange(acf_diff_plot, pacf_diff_plot, ncol = 2)
```
# 4. As observed in the ACF plot, the series exhibits strong seasonality, as evidenced by significant correlations at the 12-lag and 24-lag intervals. Therefore, seasonal differencing will be applied to address this seasonality.
```{r fig.width=14, fig.height=5}
# Seasonal differencing
fcpi_tsibble <- fcpi_tsibble %>%
  mutate(fcpi_log_seasonal_diff = c(rep(NA,12),diff(fcpi_log_diff, lag = 12)))
# Plot the seasonally differenced time series
autoplot(fcpi_tsibble, fcpi_log_seasonal_diff) +
  ggtitle("Differenced Log CPI (First Difference + Seasonal Difference)") +
  xlab("Time") + ylab("Differenced log(CPI)")
# Plot ACF and PACF
acf_seasonal_diff_plot <- ggAcf(fcpi_tsibble$fcpi_log_seasonal_diff) +
  ggtitle("ACF of Differenced Log CPI(First Difference + Seasonal Difference)") 

pacf_seasonal_diff_plot <- ggPacf(fcpi_tsibble$fcpi_log_seasonal_diff) +
  ggtitle("PACF of Differenced Log CPI(First Difference + Seasonal Difference)")

grid.arrange(acf_seasonal_diff_plot, pacf_seasonal_diff_plot, ncol = 2)
```
# 5. Now we apply ADF test and KPSS test to assess the stationary of the processed time series.
```{r}
adf_test <- ur.df(na.omit(fcpi_tsibble$fcpi_log_seasonal_diff), type = "drift", selectlags = "AIC")
summary(adf_test)
```
# ADF:
The ADF test checks if a time series is stationary. The null hypothesis is that the series has a unit root (non-stationary), and the alternative is that it is stationary.

Key Results:
Test Statistic: -14.5889, which is much smaller than the critical value at the 1% level (-3.43), so we reject the null hypothesis.

p-value: Very small (well below 0.05), confirming the rejection of the null hypothesis.

Conclusion:
Since both the test statistic and p-value suggest rejecting the null hypothesis, we conclude that the series is stationary.
```{r}
kpss_test <- kpss.test(na.omit(fcpi_tsibble$fcpi_log_seasonal_diff), null = "Level")
kpss_test
```
# KPSS:
The KPSS test checks for stationarity around a level. The null hypothesis is that the series is stationary, and the alternative is that it is non-stationary.

Test Results:
KPSS statistic: 0.12678

p-value: 0.1

At the 5% significance level, we fail to reject the null hypothesis.

Conclusion:
The series is stationary.

# 6. Next, fit SARIMA models. The PACF plot shows a clear cutoff after lag 3 in the non-seasonal component, suggesting a potential value of p = 3. In the seasonal component, significant spikes are observed at lags 12 and 24, indicating that P = 2 may be appropriate. The ACF plot exhibits a gradual decay in the non-seasonal part, and thus is not used for identifying q. However, a prominent spike at lag 12 in the seasonal ACF suggests that Q = 1 could be a suitable choice. Consequently, this initial analysis suggests that a possible model for these data is an SARIMA(3,1,0)(2,1,1)[12]. I fit this model, along with some variations on it, and compute the AIC, BIC and AICc values.

```{r}
# Fit several SARIMA models


model_result <- fcpi_tsibble %>%
  model(
    sarima_310_211 = ARIMA(fcpi_log ~ pdq(3,1,0) + PDQ(2,1,1,period=12)+0),
    sarima_311_211 = ARIMA(fcpi_log ~ pdq(3,1,1) + PDQ(2,1,1,period=12)+0),
    sarima_312_211 = ARIMA(fcpi_log ~ pdq(3,1,2) + PDQ(2,1,1,period=12)+0),
    sarima_310_111 = ARIMA(fcpi_log ~ pdq(3,1,0) + PDQ(1,1,1,period=12)+0),
    sarima_311_111 = ARIMA(fcpi_log ~ pdq(3,1,1) + PDQ(1,1,1,period=12)+0),
    sarima_312_111 = ARIMA(fcpi_log ~ pdq(3,1,2) + PDQ(1,1,1,period=12)+0),
    sarima_310_011 = ARIMA(fcpi_log ~ pdq(3,1,0) + PDQ(0,1,1,period=12)+0),
    sarima_311_011 = ARIMA(fcpi_log ~ pdq(3,1,1) + PDQ(0,1,1,period=12)+0),
    sarima_312_011 = ARIMA(fcpi_log ~ pdq(3,1,2) + PDQ(0,1,1,period=12)+0),
    sarima_210_211 = ARIMA(fcpi_log ~ pdq(2,1,0) + PDQ(2,1,1,period=12)+0),
    sarima_211_211 = ARIMA(fcpi_log ~ pdq(2,1,1) + PDQ(2,1,1,period=12)+0),
    sarima_212_211 = ARIMA(fcpi_log ~ pdq(2,1,2) + PDQ(2,1,1,period=12)+0),
    sarima_210_111 = ARIMA(fcpi_log ~ pdq(2,1,0) + PDQ(1,1,1,period=12)+0),
    sarima_211_111 = ARIMA(fcpi_log ~ pdq(2,1,1) + PDQ(1,1,1,period=12)+0),
    sarima_212_111 = ARIMA(fcpi_log ~ pdq(2,1,2) + PDQ(1,1,1,period=12)+0),
    sarima_210_011 = ARIMA(fcpi_log ~ pdq(2,1,0) + PDQ(0,1,1,period=12)+0),
    sarima_211_011 = ARIMA(fcpi_log ~ pdq(2,1,1) + PDQ(0,1,1,period=12)+0),
    sarima_212_011 = ARIMA(fcpi_log ~ pdq(2,1,2) + PDQ(0,1,1,period=12)+0),
  )

# Compare the models
glance(model_result)
```
# 7. Among all the candidate models, the SARIMA(2,1,1)(0,1,1)[12] model exhibited the lowest values for AIC, BIC, and AICc, indicating superior model performance. Consequently, it was chosen as the optimal model. The following is the model diagnostics for SARIMA(2,1,1)(0,1,1)[12].

```{r fig.width=14, fig.height=8}

# Select the best model from the list
best_sarima_model <- sarima_models %>% select(sarima_211_011)

# Generate residual diagnostic plots
gg_tsresiduals(best_sarima_model)


```
Based on residual diagnostic:

-   The residuals fluctuate randomly around zero, this means there is no trend or pattern left in the residuals.
-   The ACF plot shows all autocorrelations are inside the confidence bands, this suggests there is no significant autocorrelation in the residuals.
-   The histogram of the residuals looks like a normal distribution, this indicates the residuals are approximately normally distributed.
-   There is no sign of seasonality or structure in the residual plot,  this means the model has captured the seasonal and trend components well.

Conclusion: The residuals behave like white noise. This confirms that SARIMA(2,1,1)(0,1,1)[12] is a good model and fits the CPI data well.

# 8. Ljung Box Test.

```{r}
augment(best_sarima_model) %>%
  features(.innov, ljung_box)
```
# 9. Since the p-value is greater than 0.05, we fail to reject the null hypothesis. This means that the residuals from the SARIMA(2,1,1)(0,1,1)[12] model do not exhibit significant autocorrelation. In the end, the following plots show the 36-month forecast generated by the selected SARIMA model, with one plot representing the logged data and the other showing the original data.
```{r fig.width=10, fig.height=6 }

# Generate 36-month forecast
sarima_forecast <- best_sarima_model %>%
  forecast(h = "36 months")

# Convert forecast to tibble
forecast_tbl <- sarima_forecast %>% as_tibble()

# Extract forecast interval (80% & 95%)
forecast_interval <- sarima_forecast %>%
  hilo(level = c(80, 95)) %>%
  unpack_hilo(`80%`, names_sep = "_") %>%
  unpack_hilo(`95%`, names_sep = "_")

# Plot forecast with uncertainty intervals
autoplot(sarima_forecast, fcpi_tsibble) +
  ggtitle("Forecast from SARIMA(2,1,1)(0,1,1)[12] Model") +
  xlab("Year") + ylab("Log CPI Forecast") +
  theme_light()

# Convert log CPI forecast to original CPI scale
forecast_full_original <- forecast_interval %>%
  mutate(
    mean_original = exp(.mean),
    lower_80_original = exp(`80%_lower`),
    upper_80_original = exp(`80%_upper`),
    lower_95_original = exp(`95%_lower`),
    upper_95_original = exp(`95%_upper`)
  )

forecast_full_original

# Plot forecast in Original Data
ggplot() +
  # 95% prediction interval
  geom_ribbon(data = forecast_full_original,
              aes(x = time, ymin = lower_95_original, ymax = upper_95_original),
              fill = "lightblue", alpha = 0.5) +
  # 80% prediction interval
  geom_ribbon(data = forecast_full_original,
              aes(x = time, ymin = lower_80_original, ymax = upper_80_original),
              fill = "blue", alpha = 0.5) +
  # Forecast mean (original CPI)
  geom_line(data = forecast_full_original,
            aes(x = time, y = mean_original), color = "blue") +
  # Historical data (original CPI)
  geom_line(data = fcpi_tsibble,
            aes(x = time, y = cpi), color = "black") +
  labs(
    title = "Forecast from SARIMA Model (Original CPI Scale)",
    x = "Year",
    y = "CPI"
  ) +
  theme_light()

```